---
title: "EHR-Data-Example"
output: html_document
---

##Load libraries we need
```{r}
# install.packages("readr")
# install.packages("tidyverse")
# install.packages("assertr")

library(readr)
library(tidyverse)
library(assertr)
```


## Read in the data and verify the data has the variables we know must be there (odi_percent_1yr and....)
```{r read in main data set}
# Read in full dataset
lumbar_df <- 

# Look at the data

```

##Check format of fake_id variable
```{r}

```


## Make sure id column is in correct format
```{r}
lumbar_df_correct_id <- 

# Check that it worked
```

##Read in demographic data and verify it has required variables
```{r read in additional demographic data}

# Read in the demographic data
lumbar_age_weight_height <- 

# Look at data

```

##Rearrange columns so id is first
```{r}

```


##Join the two datasets together so that only rows in both data frames are kept
```{r}
# Only join the two datasets that have a matching fake_id value in both

```

Who did we drop? Let's investigate further.
```{r}
# Who did we drop?

```

##Fix the broken ids
```{r}
lumbar_df_new_id <- 

# Did it work?

```

##Join the datasets together, but what key do we use?
```{r}
# Throws an error
inner_join(lumbar_df_new_id, lumbar_age_weight_height, by = "?????")
```

Tell the join what variables are the same
```{r}
# join the datasets together
lumbar_full_df <- inner_join(lumbar_df_new_id, lumbar_age_weight_height, by = "??????")
```

## Determine what variables are in the dataset
```{r}
# Look at names

```

Better way to look at variables
```{r}
# Better way to look at the names

```

## Data manipulation
1. Drop all cpt codes
2. Drop variables with missing values?
```{r}
# Get rid of cpt codes
lumbar_no_cpt <- 

# Did it work?

```

# drop all missing values from our new dataset
```{r}
lumbar_no_cpt %>% drop_na()
```

What happened? 

##Explore missing with data explorer
```{r}
# install.packages("DataExplorer")
library(DataExplorer)

```

What variables are causing the most missingness? Let's only keep variables that only have less than 5% missingness. 

```{r}
# Useful for dealing with missing values
# install.packages("naniar")
library(naniar)

lumbar_no_miss <- 

# Investigate again

```

##What variables did we just drop
```{r}
# What variables did we drop?
setdiff(names(lumbar_no_cpt), names(lumbar_no_miss))
```

## Select only variables we're interested in

Variables:

odi_percent_1yr
gender
race
age
diabetes
smoker
employment
height
weight
median_household_income
asthma

```{r}
# Variables we want to select
variables_to_select <- c("odi_percent_1yr",
                         "odi_percent_preop",
                         "gender", 
                         "race", 
                         "age", 
                         "diabetes",
                         "smoker", 
                         "employment", 
                         "height", 
                         "weight",
                         "median_household_income",
                         "asthma")

# Create new data set with just the variables we want
lumbar_final <- 
```

## Explore variables with Data Explorer: Boxplots (outcome we care about is odi_percent_1yr)

```{r}
# Boxplots

```

## Explore variables with Data Explorer: Barplots
```{r}
# Barplots

```

## How many observations do we have in race and employment? (Make a table)
```{r}
# install.packages("janitor")
library(janitor)

# table for race

```

# Only keep 2 categories for race and create another table to ensure it worked
```{r}
lumbar_final <- 


```

## Check employment variable (make table)
```{r}

```

## Only keep if there are at least 200 observationsin the category for employment
```{r}
lumbar_final <- 

# Did it work?
lumbar_final %>% tabyl(employment_lumped)
```

## Create change in ODI percent (name it odi_change) and view it in descending order 
```{r}
lumbar_final <- 

# Check that it worked and vriew in descending order

```

##Does smoking have an affect, on average, for change in ODI score?
```{r}
# difference among smoking groups

```


## Do men or women who smoke have a larger change in ODI score, on average? 
```{r}
# By gender

```

##Within men and women, are there differences (on average) between different smoking groups?
```{r}
# By gender and smoking status

```


## Graphically represent the spread of the change among men, women, and smokers
```{r}
# First pass

```

##What values do we need to be able to visualize on the y-axis?
```{r}
# What are the maximum and minimum values we need to be able to see?

```

##Remake the graph
```{r}
# change the y-axis so we can see it

```

##What is the shape of the change in odi score

```{r}
# First pass

```

##Do we actually have data at all of those values? Is the default bin size meaningful?
Let's remake the graph increasing the number of bins. But how many should we choose?
```{r}
# determine the number of unique values for odi_change

```

##Remake our histogram
```{r}
# Change the number of bins

```


##Create a new variable based on both smoking status and diabetes. 
If people were previous smokers and had diabetes call it "previous_yes"
If people are current smokers and have diabetes call it "current_yes"
If people do not smoke and have diabetes call it "no_yes"
If people do not smoke and do not have diabetes call it "no_no"
```{r}

```

##Why NA values?
Now do the same as above, but for any category we didn't include, create a value that is "dont_care"
```{r}

```

##Faster way to do this if we want all combinations?

```{r}
lumbar_final <- 


```

##How about a 3-way interaction between smoker, diabetes, and race?
```{r}
# Create 3-way interaction among race, diabetes, and smoker and look at all unique values

```


## Get the mean of every column that is numeric data
```{r}

```

## Graphically, represent how preoperative and postoperative odi scores relate to smoking status and diabetes
```{r}

```

#Second attempt at above graph
```{r}

```

